---
title: "MEC4 - Lobau PARAFAC modelling"
author: "Angela Cukusic"
output:
  html_document:
    toc: true               # Table of contents
    toc_float: true         # Keep the table of contents floating on the page
    code_folding: hide      # Allow code folding in HTML output
    code_download: true     # Enable code download button
    code_copy: true         # Enable code copy button
    number_sections: true   # Numbering the sections
    theme: united           # Choose a Bootstrap theme (check available themes)
    highlight: tango        # Syntax highlighting style (pick a style from available themes)


  

---



# package loading
```{r, warning=FALSE, results='hide'}
library(staRdom)
library(tidyverse)
library(readxl)
library(vegan)
library("FactoMineR")
library("factoextra") 

```

# Fluorescence data




We should throughout all these steps be using the command: 
"eem_overview_plot(nameX..., spp = ..., contour = TRUE)"
to check between the steps the intermediate results, but since it is intensive,
I do not do it for the markdown.




## Data import
Indicate path to data and number of cores for parallel processing.
```{r, warning=FALSE, results='hold'}
# path to aqualog .dat files
sample_dir <- ("C:/Users/Angela Cukusic/Documents/faks/Ecology and Ecosystems/MEC-4 (Intro to research - s prof)/Data/flo_data") 

# number of cores
cores <- detectCores(logical = FALSE)

# import data
eem_list <- eem_read(sample_dir, import_function = "aqualog", recursive = TRUE)

```


## Process and correct
Subtract blank from data (skip if already done in aqualog software, 
but in our machine that is not the case).
```{r, warning=FALSE, results='hold'}
eem_list_blank_subtracted <- eem_extend2largest(eem_list, interpolation = 1, 
                                                extend = FALSE, cores = cores)

eem_list_blank_subtracted <- eem_remove_blank(eem_list_blank_subtracted)

# 12 blanks averaged -> checks out since we did one full year of data
```

Normalize data to Raman Units relative to blank.
```{r, warning=FALSE, results='hold'}
# if blank subtraction done in previous step
eem_list_rel_ru <- eem_raman_normalisation2(eem_list_blank_subtracted, blank = "blank")

```

At this point blanks are not needed anymore and can be removed from the sample list.
```{r, warning=FALSE, results='hold'}
eem_list_rel_ru <- eem_extract(eem_list_rel_ru, 
                               c("nano", "miliq", "milliq", "mq", "blank"), 
                               ignore_case = TRUE)
```



Remove Raman and Rayleigh scatter; 
scatter width might need fine tuning to sufficiently mask scatter.
```{r, warning=FALSE, results='hold'}
eem_list_scatter_removed_1 <- eem_rem_scat(eem_list_rel_ru, 
                                           remove_scatter = c(TRUE, TRUE, TRUE, TRUE),
                                           remove_scatter_width = c(20, 15, 25, 15))

# plot (go back and re-adjust scatter width if necessary)
#eem_overview_plot(eem_list_scatter_removed_1, spp = 7, contour = TRUE)

# or to plot each sample by one:
# ggeem(eem_list_scatter_removed_1, contour = TRUE, fill_max = 1, spp=7)
```

Interpolate missing values after removing scatter.
```{r, warning=FALSE, results='hold'}
eem_list_interp_1 <-  eem_interp(eem_list_scatter_removed_1, 
                                 cores = cores, type = 1, 
                                 extend = FALSE)

#summary(eem_list_interp_1)
```

Smooth data for peak picking.
```{r, warning=FALSE, results='hold'}
eem4peaks <- eem_smooth(eem_list_interp_1, n = 4, cores = cores)
```

## Peak picking
Calculate Coble peaks and fluorescence indices.
```{r, warning=FALSE, results='hold'}
eem_indices_and_peaks <- as.data.frame(cbind(eem_biological_index(eem4peaks),
                                 hix = eem_humification_index(eem4peaks)[, 2],
                                 fi = eem_fluorescence_index(eem4peaks)[, 2],
                                 eem_coble_peaks(eem4peaks)[, -1]))

colnames(eem_indices_and_peaks) <- c(# keep the 1:4 colnames as they are
                                     colnames(eem_indices_and_peaks)[1:4],
                                     # add to b, t, a, m, c the prefix "coble_peak"
                                     paste("coble_peak", colnames(eem_indices_and_peaks)[5:9], 
                                           sep = "_"))



```


## PARAFAC modelling

As before, apply spectral correction to remove scatter. But this time set the range even before starting, because we have too much scatter outside the 300-500 area: indicate range of EX and EM wavelengths to keep (everything above and below will be cut); values for ex and em might need some fine tuning.

```{r, warning=FALSE, results='hold'}
# cut data outside of indicated range
eem_list_trimmed <- eem_list_rel_ru %>% eem_range(ex = c(260, Inf), em = c(300,560))

# remove scatter
eem_list_scatter_removed_2 <- eem_rem_scat(eem_list_trimmed, 
                                           remove_scatter = c(TRUE, TRUE, TRUE, TRUE),
                                           remove_scatter_width = c(20, 15, 25, 15))

# continuosly go back and re-adjust ex if necessary, again I am saving the comp. power here
#eem_overview_plot(eem_list_scatter_removed_2, spp = 7, contour = TRUE)

```

Remove  other remnant noise (in this example at EX 280-295 nm; again fine tune as needed).
You can see the wobbly lines in the plot around this area so our goal is to remove that and interpolate the "straight" lines from the rest of the model.

```{r, warning=FALSE, results='hold'}

# with these just set the scatter areas as NAs
eem_list_scatter_removed_2 <- eem_setNA(eem_list_scatter_removed_2, 
                                        ex = 285:300, 
                                        interpolate = FALSE)

# with this interpolate the areas for smooth modeling
eem_list_interp_2 <-  eem_interp(eem_list_scatter_removed_2, 
                                 cores = cores, type = 1, 
                                 extend = FALSE)

```

Still there is too much variation in emission values of some samples. It is most visible in October, which rather than being fixed should be thrown out alltogether.
```{r, warning=FALSE, results='hold'}


# with this identify the order number of October samples which show great variation 
eem_list_scatter_removed_2 

# its 3, 15, 27, 39, 51, 63, 75

# identify in order to remove at emission wavelengths 319.645-330.957 nm which are the [13:20] emission interval

# make them NAs
eem_list_scatter_removed_2[64][[1]]$x[13:20, ] <- NA
eem_list_scatter_removed_2[65][[1]]$x[13:20, ] <- NA
eem_list_scatter_removed_2[66][[1]]$x[13:20, ] <- NA
eem_list_scatter_removed_2[67][[1]]$x[13:20, ] <- NA
eem_list_scatter_removed_2[68][[1]]$x[13:20, ] <- NA
eem_list_scatter_removed_2[69][[1]]$x[13:20, ] <- NA
eem_list_scatter_removed_2[70][[1]]$x[13:20, ] <- NA

#interpolate the NAs
eem_list_interp_2 <-  eem_interp(eem_list_scatter_removed_2, 
                                 cores = cores, type = 1, 
                                 extend = FALSE)

#eem_overview_plot(eem_list_interp_2, spp = 7, contour = TRUE)
```

### Finding component number

Set parameters.
```{r, warning=FALSE, results='hold'}
# range of components to look for 

## minimum number of components 
dim_min <- 3
## maximum number of components
dim_max <- 7

# will produce models with 3, 4, 5, 6, or 7 components, respectively



# number of random initialization from which the best model will be chosen
nstart <- 50
# maximum number of iterations
maxit <- 5000
# tolerance in parafac analysis (tolerated change in R2)
ctol <- 10^-6
```



Calculate models (using constraints of non-negative fluorescence).
```{r, warning=FALSE, results='hold'}
pf_screen <- eem_parafac(eem_list_interp_2, 
                            # how many components in model
                         comps = seq(dim_min, dim_max),
                         normalise = FALSE, 
                            # we cannot have negative em/ex data
                         const = c("nonneg", "nonneg", "nonneg"), 
                         maxit = maxit, 
                         nstart = nstart, 
                         ctol = ctol, 
                         cores = cores)

# rescale data such that maximum of each component is 1. 
pf_screen <- lapply(pf_screen, eempf_rescaleBC, newscale = "Fmax")
```



Check fit, peaks, and spectra of model components
```{r, warning=FALSE, results='hold'}
eempf_compare(pf_screen, contour = TRUE)
```

In this case, model3 with 5 components appears to give the best results 
Check for correlations between components in model3 (there should be no strong correlations).
```{r, warning=FALSE, results='hold'}

eempf_corplot(pf_screen[[3]]) # comps 1 and 2  seem highly correlated
```

### Normalization
Component 1 and 2 seem highly correlated; re-run the models with normalized data.
```{r, warning=FALSE, results='hold'}
# re-run
pf_screen_norm <- eem_parafac(eem_list_interp_2, 
                              comps = seq(dim_min, dim_max), 
                              # with this we normalize the data, as opposed to before
                              normalise = TRUE, 
                              const = c("nonneg", "nonneg", "nonneg"), 
                              maxit = maxit, 
                              nstart = nstart, 
                              ctol = ctol, 
                              cores = cores)

# rescale
pf_screen_norm <- lapply(pf_screen_norm, eempf_rescaleBC, newscale = "Fmax")


# check fit, peaks, and spectra
eempf_compare(pf_screen_norm, contour = TRUE)
```

After normalization, model3 and model4 with 4 and 5 components, respectively, appear reasonable; check for correlations in both.
```{r, warning=FALSE, results='hold'}
# model2
eempf_corplot(pf_screen_norm[[3]]) 

# model3
#eempf_corplot(pf_screen_norm[[4]]) 
```

### Screen outlier samples 
There is possibility that certain samples have too much leverage in the model and 
would be better for the model to remove them.
Leverage ranges between 0 and 1 with very atypical wavelengths/samples having a leverage near 1.

Leverage should be roughly similar across wavelengths/samples. 
If points with high leverage are identified, those should be removed and the model should be re-calculated.
```{r, warning=FALSE, results='hold'}
# calculate leverage
cpl <- eempf_leverage(pf_screen_norm[[3]])

# plot
eempf_leverage_plot(cpl, qlabel=0.1)

# none of the samples have leverage bigger than 0.3 and I will not remove any 
#####(exclude <- eempf_leverage_ident(cpl,qlabel=0.1))


# however! at ex 330 there is a huge spike in moodels with 5+ componennts!
# does not seem to be such a big problem with 4 components
```



### Check residuals
Residuals should be random across samples and should not show major peaks or troughs 
(small peaks along the diagonal are acceptable and may be ignored; mainly represent merely remnant scatter).
```{r, warning=FALSE, eval=FALSE}
eempf_residuals_plot(pf_screen_norm[[3]], 
                     eem_list_interp_2, 
                     residuals_only = TRUE, 
                     spp = 7, 
                     cores = cores, 
                     contour = TRUE)
```





## Final model 
Now that we are sure that all outliers are removed from the model, and that residuals are acceptable for this model we calculate a new model with increased accuracy. 


I do not include this output in pdf because of the time needed.
```{r, eval=FALSE}
# decrease tolerance in R deviation
ctol <- 10^-8 
# increase number of random starts
nstart = 60 
# increase number of maximum interations
maxit = 10000 
# number of suitable components identified previously
comps = 5 



# calculate model with n components
pf_n_components_1 <- eem_parafac(eem_list_interp_2, 
                                 comps = comps, 
                                 normalise = TRUE, 
                                 const = c("nonneg", "nonneg", "nonneg"), 
                                 maxit = maxit, 
                                 nstart = nstart, 
                                 ctol = ctol, 
                                 output = "all", 
                                 cores = cores)

# re-scale
pf_n_components_1 <- lapply(pf_n_components_1, eempf_rescaleBC, newscale = "Fmax")

```



```{r, eval=FALSE}

# check convergence and model performance
eempf_convergence(pf_n_components_1[[1]])

# plot
eempf_compare(pf_n_components_1, contour = TRUE)

# check leverage
eempf_leverage_plot(eempf_leverage(pf_n_components_1[[1]]))

# check correlations
eempf_corplot(pf_n_components_1[[1]], progress = FALSE)
#comp 2 and 3 too correlated?
```



Re-calculate a new rough model with more restarts and lower accuracy.
```{r, warning=FALSE, results='hold'}
pf_n_components_2 <- eem_parafac(eem_list_interp_2, 
                                 comps = 5, 
                                 normalise = TRUE, 
                                 const = c("nonneg", "nonneg", "nonneg"), 
                                 maxit = maxit, 
                                 nstart = 100, 
                                 ctol = min(ctol*100,0.01), 
                                 cores = cores)

```


Use the resulting model as starting point to re-calculate the high-accuracy model (nstarts can be lower here).
```{r, warning=FALSE, results='hold'}
pf_n_components_3 <- eem_parafac(eem_list_interp_2, 
                                 comps = 5, 
                                 normalise = TRUE, 
                                 const = c("nonneg", "nonneg", "nonneg"), 
                                 maxit = maxit, 
                                 nstart = 10, 
                                 ctol = ctol, 
                                 cores = cores, 
                                 Bstart = pf_n_components_2[[1]]$B, 
                                 Cstart = pf_n_components_2[[1]]$C)

pf_n_components_3 <- lapply(pf_n_components_3, eempf_rescaleBC, newscale = "Fmax")

```

Plot resulting components.
```{r, warning=FALSE, results='hold'}
eempf_comp_load_plot(pf_n_components_3[[1]], contour = TRUE)
```


## Split-half analysis
Validate final model using split-half analysis (generates Tucker’s Congruence Coefficient; value should be close to one (and not smaller than ~0.9-0.85) for good consistent model).
```{r, eval=FALSE}
# this is again way too intensive to knit in a pdf

#calculate split_half analysis
sh <- splithalf(eem_list_interp_2, 
                comps = 5, 
                normalise = TRUE, 
                rand = FALSE, 
                cores = cores, 
                nstart = nstart, 
                maxit = maxit, 
                ctol = ctol)

splithalf_plot(sh)
splithalf_tcc(sh)
```


## Visualize components

Calculate "amount" of components per sample.
```{r, warning=FALSE, results='hold'}
pf_missing_samples <- A_missing(eem_list_interp_1, pf_n_components_3[[1]])
final_comp_table <- as.data.frame(eempf4analysis(pf_missing_samples, eem4peaks))


comp_table <- final_comp_table %>% as.data.frame() %>% 
  dplyr::select("sample", "Comp.1", "Comp.2", "Comp.3", "Comp.4", "Comp.5") %>% 
  
  mutate( well_id = str_sub(sample, start = 1, end = 3),
          date = str_split(sample, "_", simplify = TRUE)[, 2],
          water_kind = case_when( str_detect(sample, "p") ~ "pumped", # for pumped
                                  str_detect(sample, "w") ~ "well", # for well
                                  TRUE ~ "surface") ) %>%              # for surface
 
   mutate(date = ymd(date))



```

Absolute concentrations
```{r, warning=FALSE, results='hold'}
comp_table %>% 
  dplyr::filter(water_kind != "well") %>% 
  mutate(date = as.factor(date),
         water_kind = as.factor(water_kind)) %>% 
  dplyr::select(!c( "well_id", "sample")) %>% 
  pivot_longer(cols = -c(date, water_kind), names_to = "comps", values_to = "concs") %>% 
  mutate(comps = as.factor(comps))     %>% 
 # find average from all three groundwater wells 
  group_by(date, comps, water_kind) %>% 
  summarise(concs = mean(concs)) %>% 
  ungroup() %>% 
  ggplot(aes(x = date, y = concs, fill = comps)) +
  geom_bar(stat = "identity") +
  facet_wrap(~ water_kind, ncol = 1) + 
  geom_text(aes(label = paste0(round(concs, 4) )), 
            position = position_stack(vjust = 0.5), size = 2.5) +
  theme(legend.position = "none",
        axis.ticks.y = element_blank(),
        axis.text.y = element_blank(),
        axis.title.y = element_blank()) +
  theme_bw() + 
  scale_fill_brewer(palette = "Set2") +
  labs(x=NULL, y=NULL) +
  scale_x_discrete(labels = c("Aug", "Sep", "Oct", "Nov", "Dec",
                              "Jan", "Feb", "Mar", "Apr", "May", 
                              "Jun", "Jul")) -> plt2

plt2

```

Relative concentrations
```{r, warning=FALSE, results='hold'}


comp_table %>% 
  mutate(date = as.factor(date),
         water_kind = as.factor(water_kind)) %>% 
  dplyr::select(!c( "well_id", "sample")) %>% 
  pivot_longer(cols = -c(date, water_kind), names_to = "comps", values_to = "concs") %>% 
  mutate(comps = as.factor(comps))     %>% 
 # find average from all three groundwater wells 
  group_by(date, comps, water_kind) %>% 
  summarise(concs = mean(concs)) %>% 
  ungroup() %>% 
 # find  relative abundances for each date in each water kind (well/pump/surf)
  group_by(date, water_kind) %>%
  mutate(rel_conc = concs / sum(concs) * 100) %>%
  ungroup() %>% 
  ggplot(aes(x = date, y = rel_conc, fill = comps)) +
  geom_bar(stat = "identity") +
  facet_wrap(~ water_kind, ncol = 1) + 
  geom_text(aes(label = paste0(round(rel_conc, 4) , "%")), 
            position = position_stack(vjust = 0.5), size = 4) +
  theme(legend.position = "none",
        axis.ticks.y = element_blank(),
        axis.text.y = element_blank(),
        axis.title.y = element_blank()) +
  theme_bw() + 
  scale_fill_brewer(palette = "Set2") +
  labs(x=NULL, y=NULL) +
  scale_x_discrete(labels = c("Aug", "Sep", "Oct", "Nov", "Dec",
                              "Jan", "Feb", "Mar", "Apr", "May", 
                              "Jun", "Jul")) -> plt1

plt1

```

```{r, warning=FALSE, results='hold'}

comp_table %>% 
  dplyr::filter(water_kind != "well") %>% 
  mutate(date = as.factor(date),
         water_kind = as.factor(water_kind)) %>% 
  dplyr::select(!c( "well_id", "sample")) %>% 
  pivot_longer(cols = -c(date, water_kind), 
               names_to = "comps", values_to = "concs") %>% 
  mutate(comps = as.factor(comps)) %>% 
  # find average from all three groundwater wells 
  group_by(date, comps, water_kind) %>% 
  summarise(concs = mean(concs)) %>% 
  ungroup() %>% 
  
  
  ggplot(aes(x = date, y = concs, color = water_kind, group =water_kind)) +
  geom_point() +
  geom_line() +
  facet_wrap(~ comps) + 
  theme(legend.position = "none",
        axis.ticks.y = element_blank(),
        axis.text.y = element_blank(),
        axis.title.y = element_blank()) +
  theme_bw() + 
  scale_fill_brewer(palette = "Set2") +
  labs(x=NULL, y=NULL) +
  scale_x_discrete(labels = c("Aug", "Sep", "Oct", "Nov", "Dec",
                              "Jan", "Feb", "Mar", "Apr", "May", 
                              "Jun", "Jul")) -> plt3

plt3
```



Get table with excitation/emission wavelengths for each component to identify excitation/emission maxima.
```{r, warning=FALSE, results='hold'}
comp_spectra <- eempf_export(pf_n_components_3[[1]])

# just look at which ex and em values is the Comp at 1, for example Comp1 has the exctation maxima (1) at 260 nm 
```





# Environmental data
```{r, warning=FALSE, results='hold'}
# load environmental data
env_data <- read_excel("C:/Users/Angela Cukusic/Documents/faks/Ecology and Ecosystems/MEC-4 (Intro to research - s prof)/Data/lobau_data_combined.xlsx")
env_data1 <- env_data  %>% 
  # temperature and oxygen are in two different columns for gw and sw
  mutate(oxygen = coalesce( oxygen_mg_L , oxygen_above_bottom_mg_L),
         temp = coalesce( temp_C , temp_above_bottom_C) ) %>% 
  # remove unneccessary columns
  dplyr::select(!c("well_id", "date_mm_dd_yyyy", "sample_type", 
                              "atp_total_pM", "oxygen_below_gw_surface_mg_L", 
                              "temp_below_gw_surface_C", "delta_13ch4_permille",
                              "sd_delta_13ch4", "oxygen_mg_L" , 
                              "oxygen_above_bottom_mg_L",
                              "temp_C" , "temp_above_bottom_C")) %>% 
  left_join(comp_table, ., by= c("sample" = "sample_id")) %>% 
  # remove data on depth since it would remove all the surface water entries
  select(! matches("depth|table") ) %>% 
  # few moths do not have measured parameters, remove those
  na.omit()

 rownames(env_data1) <- env_data1$sample 

# load amino acid data

```

Peform PCA on environmental variables to see if something can be simplified.
It would be good to this considering there is almost more explanatory variables than our samples which provides an over-fitted model and gives no coherent results.

```{r, warning=FALSE, results='hold'}

library(caret)

# 1. Calculate the correlation matrix to first simplify the highly correlated ones
cor_matrix <- cor(env_data1[,-c(1:9)])

# Find highly correlated variables with a threshold of 0.9
highly_correlated <- findCorrelation(cor_matrix, cutoff = 0.8)

# Remove the highly correlated variables from the dataframe
df_pca <- env_data1[,-c(1:9)][, -highly_correlated]

# Perform PCA on the preprocessed dataframe (df_pca)
pca_env <- prcomp(df_pca, scale = TRUE)


biplot(pca_env)
# if there is one sample that is way different than all the rest which I would like to remove
#tidy(pca_env) %>% filter(PC == "1") %>% filter(value > 5) %>% select("row")

# but seems not


# 2. Look at which components it makes sense to keep for further analyisis
# usually those that contribute more than 1 eigenvalue

# Get the eigenvalues of components (the st.dev)
summary(pca_env) # the first 7 components have have more than 1 
```



normal data: variance = (st.dev)^2
pca data:  eigenvalue = (st.dev)^2

"eigenvalues represent the variance explained by each principal component" -> eigen ~ variance

The first eight components explain more than 1 variance in data, but only the first three explaine more than 1% of variance so only keep PC1, PC2 and PC3 (as combinations of environmental variables) for further analysis.

Now see what contributes to each of the three PCs and if some environmental variables can be removed.
```{r, warning=FALSE, results='hold'}
library(broom)

# Extract the loadings from the PCA result and convert to tidy format
loadings_df <-   tidy(pca_env, matrix = "rotation")



# PC1
# Create the ggplot2 object for plotting the loadings
loadings_df %>% filter(PC == "1") %>% 
  #x=reorder(class,-amount,sum)
ggplot( aes(x = reorder(column, -value), y = value, group = factor(column))) +
  geom_bar(stat = "identity", position = "dodge", 
           fill= "gray90", color = "gray20") +
  geom_text(aes(label = column), position = position_dodge(width = 0.9), 
            #vjust = "inward",  
            angle = 90 , hjust = "inward") +
  labs(x = "Variables of PC1", y = "Loadings") +
  theme_minimal() +
  theme(axis.text.x = element_blank(),
        legend.position = "none")

# DOC and minerals


```

```{r, warning=FALSE, results='hold'}
# PC2 
loadings_df %>% filter(PC == "2") %>% 
  #x=reorder(class,-amount,sum)
ggplot( aes(x = reorder(column, -value), y = value, group = factor(column))) +
  geom_bar(stat = "identity", position = "dodge", 
           fill= "gray90", color = "gray20") +
  geom_text(aes(label = column), position = position_dodge(width = 0.9), 
            #vjust = "inward",  
            angle = 90 , hjust = "inward") +
  labs(x = "Variables of PC2", y = "Loadings") +
  theme_minimal() +
  theme(axis.text.x = element_blank(),
        legend.position = "none")


```


```{r, warning=FALSE, results='hold'}
# PC3
loadings_df %>% filter(PC == "3") %>% 
  #x=reorder(class,-amount,sum)
ggplot( aes(x = reorder(column, -value), y = value, group = factor(column))) +
  geom_bar(stat = "identity", position = "dodge", 
           fill= "gray90", color = "gray20") +
  geom_text(aes(label = column), position = position_dodge(width = 0.9), 
            #vjust = "inward",  
            angle = 90 , hjust = "inward") +
  labs(x = "Variables of PC3", y = "Loadings") +
  theme_minimal() +
  theme(axis.text.x = element_blank(),
        legend.position = "none")

# PC3 consists 60%+ of co2, atp, P, N




```



seems that the variables not represented on the PC1 and PC2 are represented on PC3, should still keep all three

```{r, warning=FALSE, results='hold'}

fviz_pca_biplot(pca_env, 
                # Fill individuals by groups
                geom.ind = "point",
                pointshape = 21,
                pointsize = 2.5,
                fill.ind = env_data1$well_id,
                col.ind = "black",
                alpha.var ="contrib",
                col.var = "contrib",
                gradient.cols = "RdBu",
                repel = TRUE)+
  ggpubr::fill_palette("jco")      
```



Make constrained analysis finally.
```{r, warning=FALSE, results='hold'}
# find coordinates 
# i did not scale since they are more or less the same scale as coordinates
tidy(pca_env) %>% filter(PC %in% c("1", "2", "3", "4", "5", "6") )  %>% 
  pivot_wider(values_from = "value", names_from = "PC") %>% 
  rename("PC1" = "1", "PC2" = "2", "PC3" = "3", 
         "PC4" ="4", "PC5" = "5", "PC6" = "6") %>% 
  arrange(desc(row)) %>% 
  column_to_rownames("row")-> env.dat

env_data1[,1:6] %>% 
  #filter(sample != "D05p_20200729") %>% 
  arrange(desc(sample)) %>% 
  select(!"sample")-> sp.dat

identical(rownames(env.dat), rownames(sp.dat)) #TRUE



simpleRDA <- capscale(sp.dat ~  ., data=env.dat ,
                 distance = "euclidean")

#summary(simpleRDA)
#64%% constrained 



```
Test if any of these PCs impacts the composition of DOC components 
```{r, warning=FALSE, results='hold'}
# Test of all canonical axes
anova.model <- anova.cca(simpleRDA, by='axis', step=1000)  # cap1 is sign, cap2 also

# test the whole model
anova.model2 <- anova.cca(simpleRDA, step=1000) #is sign
RsquareAdj(simpleRDA)$adj.r.squared # and explains 60%


# test env paameters
anova.model3 <- anova(simpleRDA, step=1000, by = "term")
# PC1, PC3, PC4 and PC5 sign


# new model

simpleRDA <- capscale(sp.dat ~  PC1 + PC3 + PC4 + PC5, data=env.dat ,
                 distance = "euclidean")
```

Only PC1 is significantly explaining the impact on the component composition, and explains 40% of variation. The PC1 is a combination of DOC (as expected), ec, oxygen, K, Na, co2. The nutrent availability does not seem to be impactful for the  DOC composition of the sample.
```{r, warning=FALSE, results='hold'}
# vectors
ccavectors <- as.matrix(scores(simpleRDA, display = "bp", scaling = "sites")*2.363299) %>%
  as.data.frame()

# site coordinates
env_data1[#which(rownames(env_data1) != "D05p_20200729")
  ,"well_id"] %>% 
  as.data.frame() %>% 
  rename("water_kind" = ".") %>% 
  arrange(desc(water_kind)) -> water

site_data <- scores(simpleRDA, display = "sites") %>% 
  as.data.frame() %>% 
  cbind(., water)
 
# add components
comp <- as.matrix(scores(simpleRDA, display = "species")*2.363299 )  %>% 
  as.data.frame()

# plotting
plot_cca <- 
  site_data %>% 
  ggplot( aes(x = CAP1, y = CAP2)) +
  geom_point(aes( color= water_kind), size = 3) +
  geom_point(color = "black", shape = 21, size = 3) +  
  geom_segment(data = ccavectors, aes(x = 0, y = 0, xend = CAP1, yend = CAP2), 
               size = 1.2,
               arrow = arrow(length = unit(0.5, "cm"))) +
   geom_vline(xintercept = c(0), color = "grey70", linetype = 2) +
   geom_hline(yintercept = c(0), color = "grey70", linetype = 2) +
  geom_text(data = ccavectors, aes(x = CAP1*1.1, y = CAP2*1.5, 
                                   label = rownames(ccavectors)),
                                   #nudge_x = 0.3, nudge_y = 0.3
                                   size=6 ) +
  geom_text(data = comp, aes(x = CAP1, y=CAP2, label = rownames(comp))) + 
  theme_bw() +
  labs(x = "CAP1 [44.58 %]", y="CAP2 [1.55 %]") +
  stat_ellipse(aes(color=water_kind), size = 1, alpha = 0.5) + 
  #stat_ellipse(aes(fill = cat), geom="polygon", level=0.95, alpha=0.09) +
  labs( color = "Water \nkind", fill = "water kind") +
  
     theme(legend.background = element_blank(),
           legend.box.background = element_rect(colour = "gray")) +
  scale_color_manual(values=c("#acc7d9","#5180a2","#3e647d","#d8b365"))

plot_cca


```
Conclusion: component 4 is mostly related to the surface water, while component 1, 2 and 3 are closely related to PC1 defined environmenal variables, namely minerals. 
